Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-27 prop_letter.tex
prop_letter.tex \section{Gleason} After meet with WTS...

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
In another word, the set of all quantum probability measures is one-to-one
corresponding to the set of states so that studying the set of states
is sufficient to understand all possible phenomenon regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties extending Gleason's theorem.
If we only focus on the Born QIVPMs~$\bar{\mu}$, that is, QIVPM's
induced by a state~$\rho$ according to the Born rule, we find that
if we monotonically shrink the length of intervals, the states consistent
with $\bar{\mu}$ will be closer and closer to $\rho$, and vise visa.
However, there is a QIVPM such that no state can be consistent with
it on every projector. In another word, although the Born rule and
Gleason's theorem can be extended to consider finite precision measurement,
the problem of finite precision measurement itself is not limited
in just applying error analysis in experimental results, but it highlight
some new phenomenon regarding to the foundation of quantum machines,
and it is necessary to embedding the idea of the resources available
for realistic measurement into the formulation of quantum theory.

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.~(\ref{eq:QIVPM-constraints}), the condition really need to
be checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent No matter which case Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

@@

In the limiting case when the intervals are infinitely precise, i.e.,
$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
we can formulate Gleason's theorem in our framework as follow.

@@

\noindent If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
the difference between their Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which implies $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$
by definition. For example, Table~\ref{tab:Compute-the-QIVPM} shows
that $\rho_{0}$ is consistent with $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$
and $\rho_{1}$ is consistent with $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
both on every projectors in $\events_{C}$. Indeed, they are consistent
with each other on every projectors in $\events$ that implies $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$
by Thm.\ref{thm:Finite-precision-Gleason}. This is true because the
exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is $\frac{1}{24}$~\citep{544199}.

@@

\noindent Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a non-Born QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval is listed in the column labeled by $P$ and separated
by commas.}

>>>>

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
In other words, the set of all quantum probability measures corresponds
one-to-one to the set of states so that studying the set of states
is sufficient to understand all possible phenomenon regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties extending Gleason's theorem.
If we only focus on the Born QIVPMs~$\bar{\mu}$, that is, QIVPM's
induced by a state~$\rho$ according to the Born rule, we find that
if we monotonically shrink the length of intervals, the states consistent
with $\bar{\mu}$ will be closer and closer to $\rho$, and vice versa.
However, there is a QIVPM such that no state can be consistent with
it on every projector. In another word, although the Born rule and
Gleason's theorem can be extended to consider finite precision measurement,
the problem of finite precision measurement itself is not limited
in just applying error analysis in experimental results, but it highlights
some new phenomenon regarding to the foundation of quantum machines,
and it is necessary to embedding the idea of the resources available
for realistic measurement into the formulation of quantum theory.

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.~(\ref{eq:QIVPM-constraints}), the condition really needing
to be checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent No matter which case Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

@@

In the limiting case when the intervals are infinitely precise, i.e.,
$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
we can formulate Gleason's theorem in our framework as follows.

@@

\noindent If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
the difference between their Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which implies $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$
by definition. For example, Table~\ref{tab:Compute-the-QIVPM} shows
that $\rho_{0}$ is consistent with $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$
and $\rho_{1}$ is consistent with $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
both on every projectors in $\events_{C}$. Indeed, they are consistent
with each other on every projectors in $\events$ that implies $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$
by Thm.~\ref{thm:Finite-precision-Gleason}. This is true because
the exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is
$\frac{1}{24}$~\citep{544199}.

@@

\noindent To prove the theorem, we first need to construct a QIVPM,
and then verify that no states consistent with it on all projectors.
Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a non-Born QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval is listed in the column labeled by $P$ and separated
by commas.}
